---
title: "Convolutional Neural Network Example with tinydnn"
author: "Andrey Ogurtsov"
date: "February 25, 2017"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Convolutional Neural Network Example with tinydnn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(knitr)
library(tinydnn)
knitr::opts_chunk$set(eval = FALSE)
```

This example shows how to train Convolutional Neural Networks (CNN) with
the **tinydnn** package. We will use the [MNIST](http://yann.lecun.com/exdb/mnist/)
dataset and the LeNet architecture
([mxnet realization](http://mxnet.io/tutorials/r/mnistCompetition.html)).

# Data

```{r}
# Download data from https://www.kaggle.com/c/digit-recognizer/data
train <- read.csv("train.csv")
train_x <- as.matrix(train[, -1])
train_y <- factor(train[, 1])
rm(train)

test <- read.csv("test.csv")
test_x <- as.matrix(test)
rm(test)
```

# Network Structure

Creating network structure is very straightforward; somewhat tricky part is setting right shapes for next layer after convolution or pool one.

```{r}
net <- net_seq()
net %<%
    fc(ncol(train_x), 
       ncol(train_x), 
       act = "identity") %<%
    conv(in_width        = 28,  
         in_height       = 28, 
         window_width    = 5, 
         window_height   = 5, 
         in_channels     = 1,  # set to 1 for grayscale images
         out_channels    = 20, # number of filters
         act = "tan_h") %<%
    max_pool(in_width    = 24, # smaller images after convolution
             in_height   = 24,
             in_channels = 20, # number of filters in conv
             pool_size_x = 2,
             pool_size_y = 2,
             stride_x    = 2, 
             stride_y    = 2,
             activation  = "identity") %<%
    conv(in_width        = 12, 
         in_height       = 12, 
         window_width    = 5, 
         window_height   = 5, 
         in_channels     = 20,
         out_channels    = 50,
         act = "tan_h") %<%
    max_pool(in_width    = 8, 
             in_height   = 8,
             in_channels = 50,
             pool_size_x = 2,
             pool_size_y = 2,
             stride_x    = 2, 
             stride_y    = 2,
             activation   = "identity") %<%
    fc(800, 500, act = "tan_h") %<%
    fc(500, nlevels(train_y), act = "softmax")
```

# Network Training

Train network with `batch_size = 25` and `epochs = 4`.

```{r}
net$fit(train_x, 
        train_y, 
        batch_size = 25, 
        epochs = 4, 
        verbose = TRUE)
```

# Making Predictions

```{r}
pred_train <- net$predict(train_x[1:3000, ], type = "class")
table(pred_train, train_y[1:3000])
# pred_train   0   1   2   3   4   5   6   7   8   9
#          0 294   0   0   0   0   0   0   0   1   2
#          1   0 334   0   0   0   1   0   1   1   0
#          2   1   3 321   2   1   2   0   1   0   0
#          3   0   1   0 259   0   1   0   1   1   0
#          4   1   0   1   0 290   0   1   0   0   5
#          5   0   0   1   3   0 283   0   0   3   2
#          6   2   1   0   0   0   0 305   0   0   1
#          7   0   1   3   1   0   0   0 295   0   3
#          8   0   2   1   3   0   0   1   1 272   0
#          9   1   0   0   2   4   1   0   3   2 277

pred_test <- net$predict(test_x, type = "class")
submission <- data.frame(ImageId = 1:nrow(test_x), 
                         Label = pred_test)
write.csv(submission, 
          file = "submission.csv", 
          row.names = FALSE,  
          quote = FALSE)
```

